---
title: "Computational Modeling - Week 4 - Assignment 2 - Part 1"
author: "Riccardo Fusaroli"
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(rethinking)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)


```{r}
#Saving Data
R_All = 6
R_Correct = 3
K_All = 2
K_Correct = 2
J_All = 198
J_Correct = 160
M_All = 132
M_Correct = 66

#Define the grid
dens=1000
p_grid <- seq( from=0 , to=1 , length.out=dens) 

#Define the prior 
prior <- rep( 1 , dens ) # Flat 

```

```{r}
#GRID ESTIMATION USING RICCARDOS NUMBERS

ALL = R_All
CORRECT = R_Correct

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

#Estimated knowledge is 0.5, read from graph

```

```{r}
#Quadratic approximation
globe.qa <- map(
    alist(
        w ~ dbinom(ALL,p) ,  # binomial likelihood
        p ~ dunif(0,1)     # uniform prior
), data=list(w=CORRECT) )
# display summary of quadratic approximation
precis( globe.qa )

#The quadratic approximation also estimates 0.5
```

```{r}
#Take the sum of all datapoints in the posterior that corresponds to points in the grid above 50%
sum(posterior[p_grid>0.5])

#There is a 50% chance that Riccardo knows more than 50%

```



2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.


```{r}
#GRID ESTIMATION USING KRISTIANS NUMBERS
ALL = K_All
CORRECT = K_Correct

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

#Estimated knowledge is hard to say. Given the prior, the most likely estimation would be a 100%. But there is a lot of uncertainty. 

sum(posterior[p_grid>0.5])
#Kristian has 0.8768844 chance of being above 0.5

```

```{r}
#GRID ESTIMATION USING JOSH'S NUMBERS
ALL = J_All
CORRECT = J_Correct

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

#Estimated knowledge is about 0.8

sum(posterior[p_grid>0.5])
#Josh has a 1 chance of being above 0.5

```

```{r}
#GRID ESTIMATION USING MIKKELS NUMBERS

ALL = M_All
CORRECT = M_Correct

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

#Estimated knowledge is 0.5

# Both Mikkel and Riccardo have posteriors symmetrical around 0.5. But Mikkels distribution is more narrow - a more certain estimation of the knowledge. We are as such more certain the Mikkel is at 0.5 than Riccardo. That is in this case because we have many more datapoints for Mikkel.

#Who is best? Hard to say. Both Kristian and Josh is probably about 0.5, so they are better than Riccardo and Mikkel. Josh has high certainty of being good (about 0.8), but Kristian have a chance of being higher. Given that I dont believe it is likely to have a score of 1 - which would normally be represented in priors - I believe that Josh's results are more likely to say that he is better. In any case, the certainty is good. 



```

3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.


```{r}

#REPEAT BUT NOW WITH DIFFERENT PRIOR
#Redefine the prior 
prior2 <- dnorm(p_grid, mean = 0.8, sd = 0.2)

```

```{r}
#RICCARDO:
ALL = R_All
CORRECT = R_Correct

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior2 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data_R=data.frame(grid=p_grid,posterior=posterior,prior=prior2,likelihood=likelihood) 
ggplot(Data_R,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")


```

```{r}
#KRISTIAN
ALL = K_All
CORRECT = K_Correct

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior2 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data_K=data.frame(grid=p_grid,posterior=posterior,prior=prior2,likelihood=likelihood) 
ggplot(Data_K,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

```

```{r}
#JOSH
ALL = J_All
CORRECT = J_Correct

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior2 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data_J=data.frame(grid=p_grid,posterior=posterior,prior=prior2,likelihood=likelihood) 
ggplot(Data_J,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")


```

```{r}

#MIKKEL
ALL = M_All
CORRECT = M_Correct

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior2 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data_M=data.frame(grid=p_grid,posterior=posterior,prior=prior2,likelihood=likelihood) 
ggplot(Data_M,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")


#All of the plots get more skewed towards 0.8. 
```


4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?


```{r}
#DO THE WHOLE THING AGAIN, NOW WITH 100* ON THE DATA
#WITH UNIFORM PRIORS

#RICCARDO
ALL = R_All*100
CORRECT = R_Correct*100

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

```


```{r}
#DO THE WHOLE THING AGAIN, NOW WITH 100* ON THE DATA

#RICCARDO
ALL = K_All*100
CORRECT = K_Correct*100

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")


```


```{r}
#DO THE WHOLE THING AGAIN, NOW WITH 100* ON THE DATA

#RICCARDO
ALL = J_All*100
CORRECT = J_Correct*100

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")


```


```{r}
#DO THE WHOLE THING AGAIN, NOW WITH 100* ON THE DATA

#RICCARDO
ALL = M_All*100
CORRECT = M_Correct*100

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")


```

```{r}

#DO THE WHOLE THING AGAIN, NOW WITH 100* ON THE DATA
#WITH NORMAL PRIORS

#RICCARDO
ALL = R_All*100
CORRECT = R_Correct*100

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior2 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior2,likelihood=likelihood) 
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")


```

```{r}
#KRISTIAN
ALL = K_All*100
CORRECT = K_Correct*100

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior2 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior2,likelihood=likelihood) 
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

```

```{r}
#JOSH
ALL = J_All*100
CORRECT = J_Correct*100

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior2 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior2,likelihood=likelihood) 
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

```

```{r}
#MIKKEL
ALL = M_All*100
CORRECT = M_Correct*100

#Compute the likelihood at each value in grid 
likelihood <- dbinom( CORRECT , size=ALL , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior2 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior2,likelihood=likelihood) 
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

#Now the priors mean much less. There is also much less uncertainty for all teachers. Kristian is now clearly best - the uncertainty is almost nonexistent.
```


5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?


```{r}
#Do the same, but with a different prior. The mean should be 0.5 or only very slightly higher - unless I believe they will have counterfactual knowledge, believe in myths, and will answer more wrong than right. Then they would have less than 0.5. The std. deviation would depend on how fanatically I am sceptical of their knowledge.


```

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
#library(brms)
#d <- data.frame(
#  Correct=c(3,2,160,66),
#  Questions=c(6,2,198,132),
#  Teacher=c("RF","KT","JS","MW"))

#FlatModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("uniform(0,1)", class = "Intercept"), family = binomial)
#plot(FlatModel)
#PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"))
#plot(PositiveModel)
#SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"))
#plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?



